---
title: "Batch Creator"
author: 
 - "**Compiled by:** `r Sys.info()['user']`"
date: "Compiled on `r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{float}
  - \usepackage{placeins}
  - \usepackage{parskip}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = NA)
n_selected <- nrow(batch_table)
```

```{r helpers, include=FALSE}
marginal_pct <- function(df, cols) {
  out <- lapply(cols, function(cc) {
    x <- df[[cc]]
    tab <- sort(table(x, useNA = "no"), decreasing = TRUE)
    data.frame(
      Level = latex_escape(names(tab)),
      Count = as.integer(tab),
      Percent = sprintf("%.0f\\%%", 100 * as.numeric(tab) / sum(tab)),
      stringsAsFactors = FALSE
    )
  })
  names(out) <- cols
  out
}

latex_escape <- function(x) {
  x <- gsub("\\\\", "\\\\textbackslash{}", x)
  x <- gsub("([#$%&_{}])", "\\\\\\1", x, perl = TRUE)
  x <- gsub("~", "\\\\textasciitilde{}", x)
  x <- gsub("\\^", "\\\\textasciicircum{}", x)
  x
}

render_table <- function(df, caption = NULL, font_size = 8) {
  use_longtable <- nrow(df) > 25
  tab <- knitr::kable(
    df,
    caption = caption,
    row.names = FALSE,
    format = "latex",
    booktabs = TRUE,
    longtable = use_longtable
  )

  if (requireNamespace("kableExtra", quietly = TRUE)) {
    latex_opts <- if (use_longtable) "repeat_header" else c("scale_down", "hold_position")
    return(
      kableExtra::kable_styling(
        tab,
        latex_options = latex_opts,
        font_size = font_size
      )
    )
  }

  tab
}
```

# Batch
The following batch has been created for **`r omic`**, with **`r n_selected`** samples.

- The seed used to create this batch was **`r seed`**

```{r note}
if (!is.null(note)) cat(note)
```

# Composition

\noindent
\begin{minipage}[t]{0.48\linewidth}
\centering
{\Large\textbf{Eligible cohort composition}}\par\medskip

```{r Eligible_cohort_composition, results='asis'}
eligible_marg <- marginal_pct(eligible, strata_cols)

for (cc in names(eligible_marg)) {
  cat("{\\centering\\large\\textbf{", latex_escape(cc), "}\\par}\n\\smallskip\n", sep = "")

  tab <- knitr::kable(
    eligible_marg[[cc]],
    format = "latex",
    booktabs = TRUE,
    row.names = FALSE,
    col.names = c("", "n", "Percent"),
    escape = FALSE
  )

  cat("{\\footnotesize\\centering\n", tab, "\n\\par}\n", sep = "")
  cat("\\medskip\n")
}
```

\end{minipage}\hfill
\begin{minipage}[t]{0.48\linewidth}
\centering
{\Large\textbf{Batch composition}}\par\medskip

```{r Batch_composition, results='asis'}
batch_marg <- marginal_pct(batch, strata_cols)

for (cc in names(batch_marg)) {
  cat("{\\centering\\large\\textbf{", latex_escape(cc), "}\\par}\n\\smallskip\n", sep = "")

  tab <- knitr::kable(
    batch_marg[[cc]],
    format = "latex",
    booktabs = TRUE,
    row.names = FALSE,
    col.names = c("", "n", "Percent"),
    escape = FALSE
  )

  cat("{\\footnotesize\\centering\n", tab, "\n\\par}\n", sep = "")
  cat("\\medskip\n")
}
```
\end{minipage}

\newpage
# Selected samples

```{r Samples_in_the_batch, results='asis'}
render_table(batch_table, caption = "Samples in the batch", font_size = 8)
```

```{r biobank_picklist, results='asis'}
if (!is.null(biobank_picklist)) {
  cat("\\newpage")
  cat("\n**Tube pick list**\n\n")
  render_table(biobank_picklist, caption = "Tube ID and location (available tubes only)", font_size = 8)
}
```

```{r insufficient_tubes, results='asis'}
if (!is.null(insufficient_tubes) && nrow(insufficient_tubes) > 0) {
  cat("\\newpage")
  cat("\n**Samples with insufficient available tubes**\n\n")
  render_table(insufficient_tubes, font_size = 8)
}
```

```{r unmatched_samples, results='asis'}
if (!is.null(unmatched_samples) && nrow(unmatched_samples) > 0) {
  cat("\\newpage")
  cat("\n**Unmatched samples**\n\n")
  cat("The following selected samples could not be matched to any available tube:\n\n")
  render_table(unmatched_samples, font_size = 8)
}
```
